---
title: "HW_9 CO2"
author: "Amy Hessl"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

You will need to install and load these packages to get this .Rmd to render figures:
```{r, libraries}
library(markdown)
library(knitr)
library(kableExtra)
```

```{r}
emissions <- read.csv("data/co2_emissions_tonnes_per_person_gapminder.csv")
```

***
### Questions:

1. Are the data in a flat file format? 

Yes. A csv is a type of flat file.

2. What is the structure of the data?  Are there any variables you might need to convert prior to manipulating the data?

```{r}
str(emissions[,1:10])
```

The data are one column of country names (factor) followed by 215 years of emissions data (numeric). The early years are mostly NA. 

3. What countries had per capita emissions larger than the 95th percentile in 2014? Make a table of their names and per capita emissions. Hint: Use quantile() to find the 95% cutoff.

```{r}
p95 = quantile(emissions$X2014, 0.95, na.rm=TRUE)
polluters = emissions[(emissions$X2014 > p95),]
table = data.frame(Country = polluters$country, CO2_Per_Capita = polluters$X2014)
table %>%
  kable() %>%
  kable_styling()
```

4. Make a table comparing the per capita emissions of the US and the world's highest per capita emitter (from above)  in 1925, 1950, and 2014.

```{r}
countries = emissions[emissions$country %in% c('United States', 'Qatar'),]
table = countries[,colnames(countries) %in% c('X1925','X1950','X2014')]
colnames(table) = c('1925','1950','2014')
rownames(table) = c('USA','Qatar')
table %>%
  kable() %>%
  kable_styling()
```

