---
title: "HW_9 CO2"
author: "Caleb Malay"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

You will need to install and load these packages to get this .Rmd to render figures:
```{r, libraries}
library(markdown)
library(knitr)
library(kableExtra)
```

For this assignment, please use the file _data/co2_emissions_tonnes_per_person_gapminder.csv_
You will collaborate with a partner to complete the homework. As with the previous assignment, each student should complete the assignment first, then you should merge your answers on one student's cloned HW_9 repository.  Finally, the host student should make a pull request to submit your answers. 
Turn in a _.Rmd_ file that is well-formatted and documented.

*Some Hints on Tables in Rmd:*
For fancy styling of tables for html output from .Rmd, try the kable() function in knitr package:
```
kable(my_tab)
```

Or even fancier, use the kablextra package:
```
my_tab %>%
  kable() %>%
  kable_styling()
```


***
### Questions:

1. Are the data in a flat file format? 
    Yes, it is in flat file format because it is a csv.

2. What is the structure of the data?  Are there any variables you might need to convert prior to manipulating the data?
    The structure of the data is a dataframe.
    ```{r}
    "Countries" is a factor and the years are numeric and most of them are filled with "NA." str(co2_emissions_tonnes_per_person_gapminder)
    ```
    ```{r}
    We will need to convert the "Countries" into a character. co2_emissions_tonnes_per_person_gapminder
    $country <- as.character(co2_emissions_tonnes_per_person_gapminder$country)
```
3. What countries had per capita emissions larger than the 95th percentile in 2014? Make a table of their names and per capita emissions. Hint: Use quantile() to find the 95% cutoff.
```{r}
> Percentile95 = quantile(co2_emissions_tonnes_per_person_gapminder$`2014`, 0.95)
> Countries_2014 = co2_emissions_tonnes_per_person_gapminder[(co2_emissions_tonnes_per_person_gapminder$`2014`> Percentile95),]
> Emission_Table = data.frame(Country = Countries_2014$country, PerCap = Countries_2014$`2014`
> Emission_Table %>%
+ kable() %>%
+ kable_styling()
```

Your table should look something like this:
```{r, echo=FALSE}
cntrys <- rep("country", 5)
dat <- rep("NA", 5)
ex_tab <- data.frame("Country"=dat, "PerCapEmissions 2014"=dat)
ex_tab %>%
  kable() %>%
  kable_styling()
```

4. Make a table comparing the per capita emissions of the US and the world's highest per capita emitter (from above)  in 1925, 1950, and 2014.  
```{r}
> CountriesCompared = co2_emissions_tonnes_per_person_gapminder[co2_emissions_tonnes_per_person_gapminder$country %in% c('United States', 'Qatar'),]
> WorldTable = CountriesCompared[,colnames(CountriesCompared) %in% c('1925', '1950', '2014')]
> colnames(WorldTable) = c('1925', '1950', '2014')
> row.names(WorldTable) = c('Qatar', 'United States')
> WorldTable %>%
+ kable() %>%
+ kable_styling()
```

Your finished product might look something like this:
```{r, echo=FALSE}
dat <- rep("NA", 3)
ex_tab <- data.frame("TopCountry"=dat, "United States"=dat)
row.names(ex_tab) <- c("1925", "1950", "2014")
ex_tab %>%
  kable() %>%
  kable_styling()

```

